An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-07-20 rebuttal_reviewers.txt

First, we thank the reviewers for their thoughtful commentary on our paper to keep the quality of
NIPS top-notch.

General comments:

We have been working continuously on the paper in the last month, and have been able to address a
number of our reviewer comments. For instance, we have been able to demonstrate worst-case scenarios
for references [3] and [4] with Omega(N) lower bounds for fixed proposals using Gaussians. We have
also added rigorous, numerical error bounds based on t-statistics (from the empirical mean and
variance) asymptotics from [13]. We have added bounds on the equilibrium distribution given bounds
on transition kernel error (we will borrow this from [4]) with the appropriate assumptions.

To reviewers 1, 2, 3, and 4 regarding the Barker function:

You are correct, the logistic function (aka Barker function) is less efficient with \theta' as good
as \theta. However, the tradeoff is that using the Barker function lets us better control our error
rates, and our newest experiments show that we are obtaining orders of magnitude improvements.

To reviewers 1, 2, and 3 regarding the convolution (in)stability:

To make it more stable, we can use L2 regularization and move some of the variance from X_norm to
X_corr. i.e. widen X_corr by convolving with a Gaussian with variance sigma^2, while decreasing the
allowed variance of X_norm by sigma^2. This ensures that X_corr has positive pdf everywhere.  We can
achieve CDF error (X_norm + X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.
We are substantially reworking this part of the paper.

To reviewer 1:

TODO (Daniel: Anything else you want to address)

To reviewer 2:

Regarding the difference between ours and [BDH] sec 6.3, that is close to what we're doing, but they
don't attempt to correct for the difference between normal and logistic distributions with an
additive variable, so they are stuck with a fixed error. That allows us to run our algorithm with
arbitrarily small errors (depending on batch size of course).

TODO (Daniel: Anything else you want to address)

To reviewer 3:

Regarding your Lemma 2 comment, yes, you are correct, the CLT only applies for large n and N, and we
also need n << N. We will clarify this.

To reviewer 7:

Yes increasing the minibatch size is one of the strategies we can use to control variance. The idea
is to increase it just enough, and then to keep it fixed thereafter. We should emphasize this
distinction more, and also to emphasize that this is the least desirable tactic to use (and our
newest results show that we do not need to use it anyway).

### 2016-07-21 rebuttal_reviewers.txt

We thank the reviewers for their thoughtful commentary on our paper to keep the quality of NIPS
top-notch. We start the rebuttal with a few general comments and then respond to specific reviewers.

@@

General comments:

We have been working continuously on the paper in the last month, and have been able to address a
number of our reviewer comments. For instance, we have been able to demonstrate worst-case scenarios
for references [3] and [4] with Omega(N) lower bounds for fixed proposals using Gaussians. We have
also added rigorous, numerical error bounds based on t-statistics (from the empirical mean and
variance) asymptotics from [13]. We have added bounds on the equilibrium distribution given bounds
on transition kernel error (we will borrow this from [4]) with the appropriate assumptions. 
(This section addresses theoretical results)

To Reviewers 2, 3, and 4 regarding the Barker function:

@@

You are correct, the logistic function (aka Barker function) is less efficient with \theta' as good
as \theta. However, the tradeoff is that using the Barker function lets us better control our error
rates, and our newest experiments show that we are obtaining orders of magnitude improvements.

To Reviewers 1, 2, and 3 regarding the convolution (in)stability:

@@

To make it more stable, we can use L2 regularization and move some of the variance from X_norm to
X_corr. i.e. widen X_corr by convolving with a Gaussian with variance sigma^2, while decreasing the
allowed variance of X_norm by sigma^2. This ensures that X_corr has positive pdf everywhere.  We can
achieve CDF error (X_norm + X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.
We are substantially reworking this part of the paper.

To Reviewer 2

(TODO?)

To Reviewers 1 and 3 regarding Lemma 2:

Both of you have correctly identified our implicit assumptions about Lemma 2. Reviewer_1 is right
that we approximated it for large N, so sampling without replacement approximates sampling with
replacement.  Reviewer_3 is correct in that the CLT only applies for large n and N, and we also need
n << N. Note also that we generally assume n is in the several hundreds, which is often enough for
the CLT to hold. We will clarify this in our revisions.

To Reviewers 1 and 2:

Thank you for mentioning [BDH], "MCMC Methods for Tall Data." We added this reference as part of our
ongoing revision of the paper and will substantially address issues raised in [BDH]. We are also
addressing Reviewer 2's other references.  In addition, we are addressing Reviewer 2's concern
regarding the difference between ours and [BDH] Section 6.3. Yes, it's close to what we are doing,
but they don't attempt to correct for the difference between normal and logistic distributions with
an additive variable, so they are stuck with a fixed error.

To Reviewer 7:

We should have clarified, that we have the option of increasing minibatch sizes, and the idea is
that we can start by increasing it just enough for the variance condition to apply. Once we have
identified a suitable minibatch size, we keep it fixed thereafter. In our latest revisions, we rerun
our test cases and experiment with reducing the proposal variance or increasing the temperature. By
doing so, we can run our algorithm using arbitrarily small minibatch sizes.

### 2016-07-21 rebuttal_reviewers.txt

We thank the reviewers for their thoughtful commentary on our paper to keep the quality of NIPS
top-notch. We start the rebuttal with a few general comments and then respond to specific reviewers.

General:

We have been working continuously on the paper in the last month, and have been able to address a
number of our reviewer comments. First, we have demonstrated worst-case scenarios for references [3]
and [4] with Omega(N) lower bounds on the number of samples consumed per iteration, using simple
Gaussian examples.

In addition, Section 4 (the theoretical results) has been vastly revamped.  We have added rigorous,
numerical error bounds based on t-statistics (from the empirical mean and variance) asymptotics from
[13]. We have added bounds on the equilibrium distribution given bounds on transition kernel error
(we will borrow this from [4]) with the appropriate assumptions.

@@

To Reviewers 2, 3, and 4 regarding the Barker function:

You are correct, the logistic (Barker) function is less efficient with \theta' as good as \theta,
and we explicitly mention this in our revision. The tradeoff is that its smoothness allows it to
naturally tolerate substantial variance in its input argument. This in turn will lead to a much more
efficient test on subsets of data, and our newest experiments show we have more control over error
rates.

@@

To Reviewers 1, 2, and 3 regarding the convolution (in)stability:

To make the convolution more stable, we frame it as a least squares (with L2 regularization)
minimization problem. We move some of the variance from X_norm to X_corr, i.e. widen X_corr by
convolving with a Gaussian with variance sigma^2, while decreasing the allowed variance of X_norm by
sigma^2. This ensures that X_corr has positive pdf everywhere.  We can achieve CDF error (X_norm +
X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.  We have added an entire
new section explaining this procedure.

To Reviewer 3:

(TODO respond to experiment criticism, but I really can't defend the work as it is ...)

### 2016-08-04 nips_2016.tex
latest edits

In fact we can do better than this approximation by using a more
precise limit distribution, under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
denote the $i^{th}$ absolute moment of $X$. Bentkus et al.'s work on Edgeworth
expansions \cite{Bentkus97} equation 2.2 gives:

\begin{lemma} Let $X_1,\ldots,X_n$ be
a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3. Then
\begin{equation}\label{eq:clt-bounds}
 \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
\end{equation}
where
\begin{equation}
G_N(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x)
\end{equation}
\end{lemma}
Here the limit $G_N(x,y)$ is a normal CDF plus a correction term that
decays as $n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}
G_N(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_N(x,y)$ is a linear combination of $\Phi(x)$, $\phi(x)$ and $\phi^{''}(x)$.

@@

>>>>

@@

In fact we can do better than this approximation (showing the error decreases as $O(n^{-1})$) by using a more
precise limit distribution under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
denote the $i^{th}$ absolute moment of $X$. If Cramer's condition holds, i.e. if
\begin{equation}
  \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1
\end{equation}

Bentkus et al.'s work on Edgeworth
expansions \cite{Bentkus97} equation 2.2 gives:

\begin{lemma} Let $X_1,\ldots,X_n$ be
  a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3. If $X$ satisfies Cramer's condition, 
then
\begin{equation}\label{eq:clt-bounds}
 \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
\end{equation}
where
\begin{equation}
G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x)
\end{equation}
\end{lemma}
This lemma shows that the average of the $X_i$ has a more precise, skewed CDF limit $G_n(x,y)$
where the skew term has weight proportional to a certain measure of skew derived
from the moments: $\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric,
the weight of the correction term is zero, and the CDF of the average of the $X_i$ converges
to $\Phi(x)$ at a rate of $O(n^{-1})$.

Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that
decays as $n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where $\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}
  \label{eq:GNderivatives}
G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_n(x,y)$ is a linear combination of
$\Phi(x)$, $\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In
Algorithm 1, we correct for the difference in $\sigma$ between
$\Delta^*$ and the variance needed by $X_{\rm corr}$ using $X_{\rm
  norm}$. This same method works when we wish to estimate the error in $\Delta^*$ vs
$G_n(x,y)$. Since all of the component functions of $G_n(x,y)$ are
derivatives of a (unit variance) $\Phi(x)$, adding a normal variable
with variance $\sigma'$ increases the variance of all three functions
to $1+\sigma'$. Thus we add $X_{\rm norm}$ as per Algorithm 1
preserving the limit in equation (\ref{eq:GNderivatives}).

The deconvolution approach can be used to constuct a correction
variable $X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard
logistic function. An additional complexity is that $G_n(x,y)$ has
additional parameters $y$ and $n$. Since these act as a single
multiplier $\frac{y}{6\sqrt{n}}$ in equation
(\label{eq:GNderivatives}), its enough to consider a function
$g(x,y')$ parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can
be computed and saved offline. As we have shown above, errors in the
``limit'' function propagate directly through as errors in the
acceptance test.  To achieve a test error of say $1e-6$ (close to single
floating point precision), we need a $y'$ spacing of $1e-6$. It
should not be necessary to tabulate values all the way to $y'=1$,
since $y'$ is scaled inversely by the square root of monibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.
Since our $x$ resolution is 10,000, this leads to a table with about
1 billion values, which can comfortably be stored in memory. However,
if $g(x,y)$ is moderately smooth in $y$, it should be possible to
achieve similar accuracy with a much smaller table. We leave further
analysis and experiments with $g(x,y)$ as future work.
